---
title: "Investigating the COVID-19 Outbreak"
subtitle: "Mon April 6th"
author: "Team 2!: Delaney Demark, Jenny Huang, Abby Mapes, Harshavardhan Srijay"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height=6, fig.width = 9, fig.align = "center")
```

```{r load-packages}
library(tidyverse)
library(broom)
library(infer)
```

## Cleaning Data

```{r load-data}
covid <- read_csv("data/covid.csv")
```

```{r clean-data}
covid <- covid %>%
  mutate(reporting_date = `reporting date`, visiting_wuhan = `visiting Wuhan`, 
         from_wuhan = `from Wuhan`, 
         if_onset_approximated = If_onset_approximated)%>%
  select(id, reporting_date, location, country, gender, age, symptom_onset, if_onset_approximated, hosp_visit_date, exposure_start, exposure_end, 
         visiting_wuhan, from_wuhan, death, recovered, source)

covid <- covid %>%
  mutate(id = factor(id), if_onset_approximated = factor(if_onset_approximated), 
         visiting_wuhan = factor(visiting_wuhan), 
         from_wuhan = factor(from_wuhan), death = factor(death), recovered = factor(recovered))%>%
  mutate(reporting_date = as.Date(reporting_date, "%m/%d/%Y"), 
         symptom_onset = as.Date(symptom_onset, "%m/%d/%Y"), 
         hosp_visit_date = as.Date(hosp_visit_date, "%m/%d/%Y"), 
         exposure_start = as.Date(exposure_start, "%m/%d/%Y"), 
         exposure_end = as.Date(exposure_end, "%m/%d/%Y"),
         death = ifelse(death == 0, "No", "Yes"))
```

```{r add age group, echo = FALSE}
covid <- covid %>%
  mutate(age_group = ifelse(age < 20, "0-19", (ifelse(age <40, "20-39",
  ifelse(age < 60, "40-59", ifelse(age < 80, "60-79", "80-99"))))))%>%
  mutate(age_group = ifelse(is.na(age_group), "NA", age_group))
```

## Section 1: Introduction

Several recent reports suggest that older age groups show more severe symptoms 
in the face of COVID-19, causing the virus to be more deadly for older age 
groups. We want to test whether this is true by comparing the death rate of 
older individuals to the death rate of younger individuals. Our null 
hypothesis (H0 ) is that the death rate for older individuals is the same as 
the death rate for younger individuals, while our alternative hypothesis (H1) 
is that the death rate for older individuals is higher than that of younger 
individuals. 

We plan on working with the data from the early stages of the COVID-19 
outbreak from 1/20/2020 to 2/15/2020. This dataset, from Kaggle, was first 
extracted from information provided by Johns Hopkins University. Johns 
Hopkins University collected this data from the World Health Organization, 
the Center for Disease Control and Prevention, the European Centre for Disease 
Prevention and Control, the National Health Commission of the People’s 
Republic of China, among other state and national government health 
departments. Each observation in the dataset is a case in which an individual 
tested positive for COVID-19. The variables include the ID number of the 
individual, the number that the case is in the country, the date the case was 
reported, the location of the case, the gender of the individual, the age of 
the individual, the age group of the individual, the date of the onset of 
symptoms, the date of the hospital visit, the start and end dates of exposure 
to the virus, if the individual visited Wuhan, if the individual was from 
Wuhan, if the patient died, and if the patient recovered. 

Link to Data: https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-
dataset#COVID19_line_list_data.csv


## Section 2: Data Analysis Plan

Through our analysis, we will use death as our outcome variable by analyzing 
the proportion of patients who have died, indicated by a value of “yes” for 
death. To do so, we will use the following predictor variables: 
age_group, gender, visited_wuhan, from_wuhan, country. 

Using these variables, we will attempt to determine not only if age affects 
one’s survival outcome due to COVID-19, but also if any of these other 
characteristics are associated with one’s survival outcome. 

To start, we will preform some preliminary exploratory data analysis to learn 
more about our data.

First, we will determine the death rate, proportion of those dead, of all 
patients in our dataset. As we can see, there is a small percentage, about 6%, 
of patients who died from COVID-19 in our data set. 

```{r death-rate}
covid %>%
  mutate(death = ifelse(death == "No", "Alive", "Dead"))%>%
  count(death)%>% 
  mutate(prop_dead = n / sum(n))
```

Now, we will look at the survival rate by age group.

```{r death-rate-by-age}
covid %>%
  mutate(death = ifelse(death == "No", "Alive", "Dead")) %>%
  group_by(age_group) %>% 
  count(death)%>% 
  mutate(prop_dead = n / sum(n))
```
The death rate is lowest for the 20-39 age group
while highest for the 80-99 age group. Older age
groups, then, tend to have a higher death rate.

Now, we will visualize survival status faceted by age_group.

```{r survival-status-by-age, fig.height=5, fig.width=5}
covid %>%
  mutate(death = ifelse(death == "Yes", "Dead", "Alive")) %>%
  filter(age_group != "NA") %>% 
  ggplot(mapping = aes(x = death, fill = death)) + 
  geom_bar(stat = "count") +
  facet_wrap( ~ age_group, ncol = 3) +
  theme_bw() +
  labs(title = "Survival Status By Age", x = "Survival Status",
       y = "Number of Patients", fill = "Survival Status")
```

We will also visualize the ages of patients for each survival status.

```{r age-vs-survival-status, fig.height=5, fig.width=5}
covid %>%
  mutate(death = ifelse(death == "Yes", "Dead", "Alive"))%>%
  ggplot(mapping = aes(x = age, fill = death)) + 
  geom_histogram(binwidth = 10, color = "black") +
  facet_wrap( ~ death, ncol = 1) +
  theme_bw() +
  labs(title = "Age of Patient vs. Survival Status", x = "Age",
       y = "Number of Patients", fill = "Survival Status")
```

Below is a summary table of the mean age of patients who have died from COVID-19
and those that haven't, out of all patients in our data set that reported their 
ages. As we can see, the average age of patients that have died is greater than 
the average age of patients that are alive. These ages give us a reference 
point to determine what age will be considered "old" and what ages will be 
considered "young" for our exploratory data analysis.

```{r mean-age-summary-table}
covid %>%
  filter(!is.na(age)) %>%
  mutate(death = ifelse(death == "Yes", "Dead", "Alive"))%>%
  group_by(death)%>% 
  summarise(mean_age = mean(age))%>% 
  arrange(desc(mean_age))
```
To understand some of our other explanatory variables, we will calculate some 
statistics to get a sense of our data in terms of gender, country, and patients 
who have been to Wuhan recently. As we can see below, around 35% of patients 
are female, 48% of patients are male and 17% of patients are not classified. 
Noting that more male patients are included in our data set will be important 
and helpful when performing our exploratory data analysis.

```{r gender-distribution}
covid %>%
  count(gender)%>% 
  mutate(prop = n / sum(n))
```

Additionally, the patients in our data set come from 38 different countries.
```{r country-count}
covid %>%
  count(country)%>% 
  mutate(total_countries = nrow(.))%>%
  select(total_countries)%>%
  slice(1:1)
```

However, as we can see below, 8 countries in our dataset have at least 1 
reported death, with China having the most number of deaths by a significant
margin, as of the date of our dataset. It will be helpful to know that 
only 8 of the 38 total countries in our data set have reported deaths when we 
consider the explanatory variable 'country' in our exploratory data analysis.

```{r age-country-summary-table}
covid %>%
  count(country, death) %>% 
  group_by(country) %>% 
  mutate(prop_dead = n / sum(n)) %>%
  filter(death == "Yes")%>%
  mutate(people_dead = n)%>%
  select(country, people_dead, prop_dead)
```

Additionally, from the summary table below, we see the mean age of the patients 
who have died for each country where there are reported deaths. In China, the 
mean age is about 71. In Taiwan, the mean age is 65. In Hong Kong, the mean age 
is about 54. However, since there were less than ten recorded deaths in 
countries other than China, the use of country as a predictor may yield
misleading results, so we plan to group the categories into "China" and 
"Other Countries."

```{r age-per-country}
covid %>%
  filter(!is.na(age)) %>%
  group_by(country)%>% 
  filter(death == "Yes") %>%
  summarise(mean_death_age = mean(age))%>% 
  arrange(desc(mean_death_age))
```

From the table below, we see the majority of patients included in our 
data set are not from Wuhan, nor have they reported that they have previously 
visited Wuhan. Even though the majority of patients in our data set are not 
from or have been to Wuhan, it will be interesting to see if time in Wuhan is 
associated with one's survival outcome.

```{r wuhan-distribution}
covid %>%
  count(from_wuhan) %>% 
  mutate(prop = n / sum(n)) %>%
  select(from_wuhan, prop)

covid %>%
  count(visiting_wuhan) %>% 
  mutate(prop = n / sum(n)) %>%
  select(visiting_wuhan, prop)
```

We plan to use the following statistical methods to answer our 
research questions about whether age_group, gender, country, 
visited_wuhan, from_wuhan play a role in 
variable death rate:

a) Calculate the conditional probability of P(Death | Age) to determine if 
there are any confounding variables: gender, visited_wuhan, from_wuhan, country.
We will also keep in mind that the conclusions we draw about death rate may be 
confounded with variables that are not included in our dataset, 
such as smoking/drinking status, chronic disease status, BMI, etc.

b) Use estimation via bootstrap to create a confidence interval 
for the mean age of affected individuals:

For each age_group, we will take bootstrap samples, calculate the mean death 
rate, and obtain a 95% confidence interval for the mean death rate of that 
particular age group. If the confidence intervals for the mean death rates do 
not overlap, we can conclude that the mean death rate for the older age group 
is significantly different from the mean death rate of the younger age group. 

If there are any confoundingvariables, we will estimate the mean age 
of affected individuals faceted by the confounding variables.


```{r ci-young-age}
covid_young_age <- covid %>%
  filter(age_group == "20-39") %>%
  count(age) %>% 
  mutate(death_rate = n / sum(n))

boot_dist_young <- covid_young_age %>%
  specify(response = death_rate) %>% 
  generate(reps = 15000, type = "bootstrap") %>% 
  calculate(stat = "mean")


boot_dist_young %>%
  summarize(lower_bound = quantile(stat, 0.025),
            upper_bound = quantile(stat, 0.975))
```

We are 95% confident that the true mean death rate of the population age 20-39 
is between the range of (0.03668033 and 0.06598361).

```{r ci-old-age}
covid_old_age <- covid %>%
  filter(age_group == "80-99") %>%
  count(age) %>% 
  mutate(death_rate = n / sum(n))

boot_dist_old <- covid_old_age %>%
  specify(response = death_rate) %>% 
  generate(reps = 15000, type = "bootstrap") %>% 
  calculate(stat = "mean")


boot_dist_old %>%
  summarize(lower_bound = quantile(stat, 0.025),
            upper_bound = quantile(stat, 0.975))
```
We are 95% confident that the true mean death rate of the population 
age 80-99 is between the range of (0.04545455 and 0.1590909).

Since the two confidence intervals overlap, we cannot claim that there is
significant difference between the mean death rate of individuals in the 
80-99 age group with those in 20-39 age group. However, both the 
lower and upper end of the confidence interval for the older age group 
(0.04545455 and 0.1590909) is higher than the interval for the younger age 
group (0.03668033 and 0.06598361).

We will repeat this process by determining the role of other explanatory 
variables on death rate: gender, country, visited_wuhan, and from_wuhan.

c) To get more information than seeing whether the confidence intervals overlap, 
we will use simulation-based hypothesis testing to give the exact probability 
estimate (p-value) for the alternate hypothesis.

H0: µ(death rate of young age group) = µ(death rate of old age group)
H1: µ(death rate of young age group) < µ(death rate of old age group)

By using the bootstrap distribution, we will see where the elderly mean 
death rate lies on the null distribution for the young mean death rate. 

First, we estimate the mean death rate among the older age group through
using bootstrap estimation.

```{r null-dist-old}
old_bootstrap <- covid_old_age %>%
  specify(response = death_rate) %>%
  generate(reps = 15000, type = "bootstrap") %>%
  calculate(stat = "mean")

old_boot_mean <- old_bootstrap %>%
  summarise(mean = mean(stat)) %>% 
  pull()
old_boot_mean
```

Next, we create a null distribution for the mean death rate of the younger 
age group.

```{r null-dist-young}
young_null_dist <- covid_young_age %>%
  specify(response = death_rate) %>%
  generate(reps = 15000, type = "bootstrap") %>%
  calculate(stat = "mean")
```

By graphing the sample mean for the older age group on the null distribution 
of the younger age group, we see that the probability of observing a sample 
mean the same or greater than that of the estimated mean for the older age 
group is very low.

```{r graph-null-dist-young, fig.height = 5, fig.width = 6}
visualize(young_null_dist) +
  shade_p_value(obs_stat = old_boot_mean, direction = "greater") +
  theme_minimal(base_size = 16)
```

We confirm this by doing a hypothesis test and getting the p_value,
which comes out to be 6.666667e-05. Therefore, the probability of seeing a mean
death rate of 0.09094273 (the sample mean death rate of the older age group) 
for the younger age group is approximately zero.

```{r perform-hypothesis-test}
p_value <- covid_young_age %>%
  specify(response = death_rate) %>%
  generate(reps = 15000, type = "bootstrap") %>%
  calculate(stat = "mean") %>% 
  get_p_value(obs_stat = old_boot_mean, direction = "greater")
```

Thus, with an alpha level of 0.05, we have sufficient evidence to reject 
the null hypothesis that the mean death rate for the age group "80-99" is equal
to the mean death rate of age group "20-39." 

d) A final way we will measure whether the explanatory variables are significant 
determiners of death rate is by creating logistic regression models. 

```{r logistic-regression}
covid_age <- covid %>%
  filter(age <= 80) %>%
  count(age) %>% 
  mutate(death_rate = n / sum(n))

m_ldeath_age <- lm(log(death_rate) ~ age, data = covid_age)
m_ldeath_age %>%
  tidy() %>%
  select(term, estimate)
```

```{r r-squared-age-model}
m_ldeath_age %>% 
  glance() %>% 
  pull(r.squared)
```

```{r visualize-age-model, fig.height = 5, fig.width = 5}
ggplot(covid_age, mapping = aes(x = age, y = death_rate)) +
  geom_point() +
  geom_smooth(method='lm') +
  labs(title = "Death Rate across Age", x = "Age", y = "Log Death Rate" )
```

From this visualization, we can see that death rate peaks around 
age 60. This may be because of the small number of patients who are 
over 70.

Since the graph does not seem to follow a linear pattern, we will check whether
a linear fit is appropriate using a residual plot.

```{r residual-plot, fig.height = 5, fig.width = 5}
m_ldeath_age_aug <- augment(m_ldeath_age)
ggplot(m_ldeath_age_aug, mapping = aes(x = .fitted, y = .resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "blue", lty = 2) +
  labs(x = "Predicted death rate", y = "Residuals") + theme_minimal()
```
From seeing a parabolic pattern in the residual plot, we know that the
linearity assumption is violated, so a linear fit would not be the best model.

Created the diagnostic plot function in order to check assumptions.

```{r plot_diagnostic_function}
plot_diagnostic <- function(x) {
  
  plot_variance <- augment(x) %>% 
    ggplot(aes(x = .fitted, y = .resid)) +
    geom_point() +
    geom_hline(yintercept = 0, color = "red", lty = 2) +
    labs(x = expression(hat(y)), y = "Residual") +
    theme_minimal(base_size = 16)
  
  plot_normality <- augment(x) %>% 
    ggplot(aes(x = .resid)) +
    geom_histogram(binwidth = 50, alpha = 0.5, color = "Red") +
    labs(x = "Residual value", y = "Count") +
    theme_minimal(base_size = 16)
  
  plot_independence <- augment(lm_volume_full) %>% 
    ggplot(aes(x = 1:nrow(covid_age), y = .resid)) +
    geom_point() +
    labs(x = "Index", y = "Residual") +
    theme_minimal(base_size = 16)
    
  plot_variance / (plot_normality + plot_independence)
}
```

```{r use_plot_diagnostic}
plot_diagnostic(x = m_ldeath_age_aug)
```

We will create logistic regression models for each variable, then determine
whether each variable is significant in influencing death rates by looking at
P-value for each coefficient.

Finally, we will use AIC/BIC model selection criteria to select for a model 
with the highest adjusted R^2 using backwards elimination.


Notes in regards to failing linearity assumption:
-fit logistic regression model on the original variable. 
(response = health outcome)
-or transform the age.
-knn classification on death (quantitative predictors)

## Section 3: Data
The dimensions of our dataset are 1,085 rows by 16 columns.
```{r glimpse-data}
glimpse(covid)
```

      