---
title: "Investigating the COVID-19 Outbreak"
subtitle: ""
author: "Team 2!: Delaney Demark, Jenny Huang, Abby Mapes, Harshavardhan Srijay"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height=6, fig.width = 9, fig.align = "center")
```

```{r load-packages}
library(tidyverse)
library(broom)
library(infer)
```

## Cleaning Data

```{r load-data}
covid <- read_csv("data/covid.csv")
```

```{r clean-data}
covid <- covid %>%
  mutate(reporting_date = `reporting date`, visiting_wuhan = `visiting Wuhan`, 
         from_wuhan = `from Wuhan`, 
         if_onset_approximated = If_onset_approximated)%>%
  select(id, reporting_date, location, country, gender, age, symptom_onset, if_onset_approximated, hosp_visit_date, exposure_start, exposure_end, 
         visiting_wuhan, from_wuhan, death, recovered, source)

covid <- covid %>%
  mutate(id = factor(id), if_onset_approximated = factor(if_onset_approximated), 
         visiting_wuhan = factor(visiting_wuhan), 
         from_wuhan = factor(from_wuhan), death = factor(death), recovered = factor(recovered))%>%
  mutate(reporting_date = as.Date(reporting_date, "%m/%d/%Y"), 
         symptom_onset = as.Date(symptom_onset, "%m/%d/%Y"), 
         hosp_visit_date = as.Date(hosp_visit_date, "%m/%d/%Y"), 
         exposure_start = as.Date(exposure_start, "%m/%d/%Y"), 
         exposure_end = as.Date(exposure_end, "%m/%d/%Y"),
         death = ifelse(death == 0, "No", "Yes"))
```

```{r add age group, echo = FALSE}
covid <- covid %>%
  mutate(age_group = ifelse(age < 20, "0-19", (ifelse(age <40, "20-39",
  ifelse(age < 60, "40-59", ifelse(age < 80, "60-79", "80-99"))))))%>%
  mutate(age_group = ifelse(is.na(age_group), "NA", age_group))
```

## Section 1: Introduction

Several recent reports suggest that older age groups show more severe symptoms 
in the face of COVID-19, causing the virus to be more deadly for older age 
groups. We want to test whether this is true by comparing the death rate of 
older individuals to the death rate of younger individuals. Our null 
hypothesis (H0 ) is that the death rate for older individuals is the same as 
the death rate for younger individuals, while our alternative hypothesis (H1) 
is that the death rate for older individuals is higher than that of younger 
individuals. 

We plan on working with the data from the early stages of the COVID-19 
outbreak from 1/20/2020 to 2/15/2020. This dataset, from Kaggle, was first 
extracted from information provided by Johns Hopkins University. Johns 
Hopkins University collected this data from the World Health Organization, 
the Center for Disease Control and Prevention, the European Centre for Disease 
Prevention and Control, the National Health Commission of the People’s 
Republic of China, among other state and national government health 
departments. Each observation in the dataset is a case in which an individual 
tested positive for COVID-19. The variables include the ID number of the 
individual, the number that the case is in the country, the date the case was 
reported, the location of the case, the gender of the individual, the age of 
the individual, the age group of the individual, the date of the onset of 
symptoms, the date of the hospital visit, the start and end dates of exposure 
to the virus, if the individual visited Wuhan, if the individual was from 
Wuhan, if the patient died, and if the patient recovered. 

Link to Data: https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset#COVID19_line_list_data.csv


## Section 2: Data Analysis Plan

Through our analysis, we will use death as our outcome variable by analyzing 
the proportion of patients who have died, indicated by a value of “yes” for 
death. To do so, we will use the following predictor variables: 
age_group, gender, visited_wuhan, from_wuhan, country. 

Using these variables, we will attempt to determine not only if age affects 
one’s survival outcome due to COVID-19, but also if any of these other 
characteristics are associated with one’s survival outcome. 

To start, we will preform some preliminary exploratory data analysis to learn 
more about our data.

First, we will determine the death rate, proportion of those dead, of all 
patients in our dataset. As we can see, there is a small percentage, about 6%, 
of patients who died from COVID-19 in our data set. 
```{r death-rate}
covid %>%
  mutate(death = ifelse(death == "No", "Alive", "Dead"))%>%
  count(death)%>% 
  mutate(prop_dead = n / sum(n))
```

Now, we will visualize the ages of patients for each survival status.

```{r age-vs-survival-status, fig.height=5, fig.width=5}
covid %>%
  mutate(death = ifelse(death == "Yes", "Dead", "Alive"))%>%
  ggplot(mapping = aes(x = age, fill = death)) + 
  geom_histogram(binwidth = 10, color = "black") +
  facet_wrap( ~ death, ncol = 1) +
  theme_bw() +
  labs(title = "Age of Patient vs. Survival Status", x = "Age",
       y = "Number of Patients", fill = "Survival Status")
```

Below is a summary table of the mean age of patients who have died from COVID-19
and those that haven't, out of all patients in our data set that reported their 
ages. As we can see, the average age of patients that have died is greater than 
the average age of patients that are alive. These ages give us a reference 
point to determine what age will be considered "old" and what ages will be 
considered "young" for our exploratory data analysis.

```{r mean-age-summary-table}
covid %>%
  filter(!is.na(age)) %>%
  mutate(death = ifelse(death == "Yes", "Dead", "Alive"))%>%
  group_by(death)%>% 
  summarise(mean_age = mean(age))%>% 
  arrange(desc(mean_age))
```
To understand some of our other explanatory variables, we will calculate some 
statistics to get a sense of our data in terms of gender, country, and patients 
who have been to Wuhan recently. As we can see below, around 35% of patients 
are female, 48% of patients are male and 17% of patients are not classified. 
Noting that more male patients are included in our data set will be important 
and helpful when performing our exploratory data analysis.

```{r gender-distribution}
covid %>%
  count(gender)%>% 
  mutate(prop = n / sum(n))
```

Additionally, the patients in our data set come from 38 different countries.
```{r country-count}
covid %>%
  count(country)%>% 
  mutate(total_countries = nrow(.))%>%
  select(total_countries)%>%
  slice(1:1)
```

However, as we can see below, there are only 3 countries with patients who have 
died in our data set: China, Hong Kong, Taiwan. It will be helpful to know that 
only 3 of the 38 total countries in our data set have reported deaths when we 
consider the explanatory variable country in our exploratory data analysis.

```{r age-country-summary-table}
covid %>%
  count(country, death) %>% 
  group_by(country) %>% 
  mutate(prop_dead = n / sum(n)) %>%
  filter(death == "Yes")%>%
  mutate(people_dead = n)%>%
  select(country, people_dead, prop_dead)
```

Additionally, from the summary table below, we see the mean age of the patients 
who have died for each country where there are reported deaths. In China, the 
mean age is about 71. In Taiwan, the mean age is 65. In Hong Kong, the mean age 
is about 54.

```{r age-per-country}
covid %>%
  filter(!is.na(age)) %>%
  group_by(country)%>% 
  filter(death == "Yes") %>%
  summarise(mean_death_age = mean(age))%>% 
  arrange(desc(mean_death_age))
```

From the table below, we see the majority of patients included in our 
data set are not from Wuhan, nor have they reported that they have previously 
visited Wuhan. Even though the majority of patients in our data set are not 
from or have been to Wuhan, it will be interesting to see if time in Wuhan is 
associated with one's survival outcome.

```{r wuhan-distribution}
covid %>%
  count(from_wuhan) %>% 
  mutate(prop = n / sum(n)) %>%
  select(from_wuhan, prop)

covid %>%
  count(visiting_wuhan) %>% 
  mutate(prop = n / sum(n)) %>%
  select(visiting_wuhan, prop)
```

In our exploratory data analysis, we plan to use the following statistical 
methods to answer some of these questions:

a) Calculate the conditional probability of P(Death | Age) and determine if 
there are any confounding variables: gender, visited_wuhan, from_wuhan, country.

b) Take a bootstrap sample to estimate a confidence interval for the mean age 
of affected individuals. If there are any confounding variables, we will 
estimate the mean age of affected individuals faceted by the confounding 
variables.

c) To answer the question of whether age_group is associated with death rate, 
we will be using simulation via bootstrap. 

For each age_group, we will take bootstrap samples, calculate the mean death 
rate, and obtain a 95% confidence interval for the mean death rate of that 
particular age group. If the confidence intervals for the mean death rates do 
not overlap, we can conclude that the mean death rate for the older age group 
is significantly different from the mean death rate of the younger age group. 

```{r ci-young-age}
covid_young_age <- covid %>%
  filter(age_group == "20-39") %>%
  count(age) %>% 
  mutate(death_rate = n / sum(n))

boot_dist_young <- covid_young_age %>%
  specify(response = death_rate) %>% 
  generate(reps = 15000, type = "bootstrap") %>% 
  calculate(stat = "mean")


boot_dist_young %>%
  summarize(lower_bound = quantile(stat, 0.025),
            upper_bound = quantile(stat, 0.975))
```

We are 95% confident that the true mean death rate of the population age 20-39 
is between the range of (0.03668033 and 0.06598361).

```{r ci-old-age}
covid_old_age <- covid %>%
  filter(age_group == "80-99") %>%
  count(age) %>% 
  mutate(death_rate = n / sum(n))

boot_dist_old <- covid_old_age %>%
  specify(response = death_rate) %>% 
  generate(reps = 15000, type = "bootstrap") %>% 
  calculate(stat = "mean")


boot_dist_old %>%
  summarize(lower_bound = quantile(stat, 0.025),
            upper_bound = quantile(stat, 0.975))
```
We are 95% confident that the true mean death rate of the population 
age 80-99 is between the range of (0.04545455 and 0.1590909).

Since the two confidence intervals overlap, we cannot claim that there is
significant difference between the mean death rate of individuals in the 
80-99 age group with those in 20-39 age group. However, both the 
lower and upper end of the confidence interval for the older age group 
(0.04545455 and 0.1590909) is higher than the interval for the younger age 
group (0.03668033 and 0.06598361).

In addition to age, we will repeat this process for the other
explanatory variables: gender, country, visited_wuhan, and from_wuhan.

d) Another way we will measure whether the explanatory variable age, gender,
visited_wuhan, from_wuhan, and country are significant determiners of death 
rate is by creating logistic regression models. 

```{r logistic-regression}
covid_age <- covid %>%
  filter(age <= 80) %>%
  count(age) %>% 
  mutate(death_rate = n / sum(n))

m_ldeath_age <- lm(log(death_rate) ~ age, data = covid_age)
m_ldeath_age %>%
  tidy() %>%
  select(term, estimate)
```

```{r r-squared-age-model}
m_ldeath_age %>% 
  glance() %>% 
  pull(r.squared)
```

```{r visualize-age-model, fig.height = 5, fig.width = 5}
ggplot(covid_age, mapping = aes(x = age, y = log(death_rate))) +
  geom_point() +
  geom_smooth(method='lm') +
  labs(title = "Death Rate across Age", x = "Age", y = "Log Death Rate" )
```

From this visualization, we can see that death rate peaks around 
age 60.

After creating logistic regression models for each variable, we will determine
whether each variable is significant in influencing death rates by looking at
P-value for each coefficient.

Then, we can use AIC/BIC model selection criteria to select for a model 
with the highest adjusted R^2 using backwards elimination.

##Section 3: Data
The dimensions of our dataset are 1,085 rows by 16 columns.
```{r glimpse-data}
glimpse(covid)
```

      